### 2020-12-16 paper/authors.tex
Adds abstract

\collaboration{1000}{The LSST Dark Energy Science Collaboration (LSST DESC)}
\noaffiliation

% paper leads

\author[0000-0001-9789-9646]{Joe Zuntz}
\affiliation{Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom}

%main partticipants and text contributors  (alphabetical)
\author{David Alonso}
\noaffiliation

\author{Bela Abolfathi}
\noaffiliation

\author{Abby Bault}
\noaffiliation

\author{Jean-Eric Campagne}
\noaffiliation

\author{Stefano Cavuoti}
\noaffiliation

\author{David Kirkby}
\noaffiliation

\author[0000-0001-7956-0542]{Fran\c{c}ois Lanusse}
\noaffiliation
%\affiliation{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}

\author{Alex Malz}
\noaffiliation

\noaffiliation
\author{Andrina Nicola}
\noaffiliation

\author{Erfan Nourbakhsh}
\noaffiliation

\author{Andy Park}
\noaffiliation

\author{Anze Slosar}
\noaffiliation

\author{Angus Wright}
\noaffiliation

% minor contributions and/or text writers (alphabetical)