### 2021-10-14 paper/authors2.tex
putting things into OJA format

\author{Joe Zuntz\altaffilmark{1}, Fran\c{c}ois Lanusse\altaffilmark{2}, Alex I. Malz\altaffilmark{3}, Angus H. Wright\altaffilmark{3}, An\v{z}e Slosar\altaffilmark{4}, Bela Abolfathi\altaffilmark{5}, David Alonso\altaffilmark{6}, Abby Bault\altaffilmark{5}, Cl\'{e}cio R. Bom\altaffilmark{7}\altaffilmark{8}, Massimo Brescia\altaffilmark{9}, Adam Broussard\altaffilmark{10}, Jean-Eric Campagne\altaffilmark{11}, Stefano Cavuoti\altaffilmark{9}\altaffilmark{12}\altaffilmark{13}, Eduardo S. Cypriano\altaffilmark{14}, Bernardo M. O. Fraga\altaffilmark{7}, Eric Gawiser\altaffilmark{10}, Elizabeth J. Gonzalez\altaffilmark{7}\altaffilmark{15}\altaffilmark{16}, Dylan Green\altaffilmark{5}, Peter Hatfield\altaffilmark{6}, Kartheik Iyer\altaffilmark{17}, David Kirkby\altaffilmark{5}, Andrina Nicola\altaffilmark{18}, Erfan Nourbakhsh\altaffilmark{19}, Andy Park\altaffilmark{5}, Gabriel Teixeira\altaffilmark{7}, Katrin Heitmann\altaffilmark{20}, Eve Kovacs\altaffilmark{20}, Yao-Yuan Mao\altaffilmark{21},}
\collaboration{The LSST Dark Energy Science Collaboration, LSST DESC}

\altaffiltext{1}{Institute for Astronomy, University of Edinburgh, Edinburgh EH9 3HJ, United Kingdom}
\altaffiltext{2}{AIM, CEA, CNRS, Universit\'e Paris-Saclay, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, F-91191 Gif-sur-Yvette, France}
\altaffiltext{3}{German Centre of Cosmological Lensing, Ruhr-Universit\"{a}t, Universit\"{a}tsstra{\ss}e 150, 44801 Bochum, Germany}
\altaffiltext{4}{Brookhaven National Laboratory, Upton NY 11973, USA}
\altaffiltext{5}{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}
\altaffiltext{6}{Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
\altaffiltext{7}{Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, 22290-180 Rio de Janeiro, RJ, Brazil}
\altaffiltext{8}{Centro Federal de Educa\c{c}\~{a}o Tecnol\'{o}gica Celso Suckow da Fonseca, Rodovia M\'{a}rcio Covas, lote J2, quadra J - Itagua\'{i} (Brazil)}
\altaffiltext{9}{INAF - Osservatorio Astronomico di Capodimonte, via Moiariello 16, I-80131, Napoli Italy}
\altaffiltext{10}{Department of Physics and Astronomy, Rutgers, the State University, Piscataway, NJ 08854, USA}
\altaffiltext{11}{Universit\'e Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France}
\altaffiltext{12}{INFN - Sezione di Napoli, via Cinthia 21, I-80126 Napoli, Italy}
\altaffiltext{13}{Department of Physics ``E. Pancini'', University of Naples Federico II, via Cintia, 21, I-80126 Napoli, Italy}
\altaffiltext{14}{Universidade de S\~ao Paulo, IAG, Rua do Mat\~ao 1225, S\~ao Paulo, SP, Brazil}
\altaffiltext{15}{Instituto de Astronom\'{\i}a Te\'orica y Experimental (IATE-CONICET), Laprida 854, X5000BGR, C\'ordoba, Argentina.}
\altaffiltext{16}{Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Laprida 854, X5000BGR, C\'ordoba, Argentina.}
\altaffiltext{17}{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St George St, Toronto, ON M5S 3H4, Canada}
\altaffiltext{18}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
\altaffiltext{19}{Department of Physics and Astronomy, University of California, Davis, Davis, CA 95616, USA}
\altaffiltext{20}{Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA}
\altaffiltext{21}{Department of Physics and Astronomy, Rutgers University, Piscataway, NJ 08854, USA}