### 2021-10-15 paper/authors2.tex

\author{Joe Zuntz\altaffilmark{1}, Fran\c{c}ois Lanusse\altaffilmark{2}, Alex I. Malz\altaffilmark{3}, Angus H. Wright\altaffilmark{3}, An\v{z}e Slosar\altaffilmark{4}, Bela Abolfathi\altaffilmark{5}, David Alonso\altaffilmark{6}, Abby Bault\altaffilmark{5}, Cl\'{e}cio R. Bom\altaffilmark{7}\altaffilmark{8}, Massimo Brescia\altaffilmark{9}, Adam Broussard\altaffilmark{10}, Jean-Eric Campagne\altaffilmark{11}, Stefano Cavuoti\altaffilmark{9}\altaffilmark{12}\altaffilmark{13}, Eduardo S. Cypriano\altaffilmark{14}, Bernardo M. O. Fraga\altaffilmark{7}, Eric Gawiser\altaffilmark{10}, Elizabeth J. Gonzalez\altaffilmark{7}\altaffilmark{15}\altaffilmark{16}, Dylan Green\altaffilmark{5}, Peter Hatfield\altaffilmark{6}, Kartheik Iyer\altaffilmark{17}, David Kirkby\altaffilmark{5}, Andrina Nicola\altaffilmark{18}, Erfan Nourbakhsh\altaffilmark{19}, Andy Park\altaffilmark{5}, Gabriel Teixeira\altaffilmark{7}, Katrin Heitmann\altaffilmark{20}, Eve Kovacs\altaffilmark{20}, Yao-Yuan Mao\altaffilmark{21},}
\collaboration{The LSST Dark Energy Science Collaboration, LSST DESC}